---
title: "607 Final Project"
author: "Weijian Cai, Qixing Li"
date: "May 6, 2019"
output: 
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE)
```

#1. Data

##1.1 Data Source

We obtained data from [Wall Street Journal](http://online.wsj.com/public/resources/documents/info-Degrees_that_Pay_you_Back-sort.html?mod=article_inline)
and from [U.S. DEPARTMENT OF EDUCATION](https://collegescorecard.ed.gov/data/).


```{r}
#Load packages and data
library(tidyverse)
library(DT)
library(jsonlite)
library(Rmisc)
```


##1.2 Data Load
Read and parse JSON format files from U.S. DEPARTMENT OF EDUCATION
```{r}
df_api = NULL
for(i in 0:100) {
  url_api <-paste0("https://api.data.gov/ed/collegescorecard/v1/schools?fields=school.name,latest.earnings.6_yrs_after_entry.mean_earnings.lowest_tercile,latest.earnings.6_yrs_after_entry.mean_earnings.middle_tercile,latest.earnings.6_yrs_after_entry.mean_earnings.highest_tercile&page=",
                   i,"&api_key=gHHgffzGEVcJZU17NEZzdgkCJ50xqfJ0y3S0gQBH")
  ed_api <- fromJSON(url_api)
  df_api <- rbind(df_api,ed_api$results)
}
```

We read the WSJ data from the CSV file which were uploaded to Github.
```{r}
# salaries by degree
url_deg <- "https://raw.githubusercontent.com/qixing810/CUNYSPS-DataScience/master/DS607/Project/final%20project/degrees-that-pay-back.csv"
df_deg <- read_csv(url_deg,
                   col_names = c("major", "start_med_slry", "mid_car_slry",
                                 "percent_chng", "mid_car_10th", "mid_car_25th",
                                 "mid_car_75th", "mid_car_90th"),  # rename the column names
                   skip = 1)  # names specified, skip header

# salaries by college type
url_col <- "https://raw.githubusercontent.com/qixing810/CUNYSPS-DataScience/master/DS607/Project/final%20project/salaries-by-college-type.csv"
df_col <- read_csv(url_col,
                   col_names = c("school_name", "school_type", "start_med_slry",
                                 "mid_car_slry", "mid_car_10th", "mid_car_25th",
                                 "mid_car_75th", "mid_car_90th"),
                   skip = 1)

# salaries by region
url_reg <- "https://raw.githubusercontent.com/qixing810/CUNYSPS-DataScience/master/DS607/Project/final%20project/salaries-by-region.csv"
df_reg <- read_csv(url_reg,
                   col_names = c("school_name", "region", "start_med_slry",
                                 "mid_car_slry", "mid_car_10th", "mid_car_25th",
                                 "mid_car_75th", "mid_car_90th"),
                   skip = 1)
```


#2. Tidy and Wrangle

##2.1 Data Structure
First we looked at the structure of the dataset.
```{r}
glimpse(df_api)
glimpse(df_deg)
glimpse(df_col)
glimpse(df_reg)
```

##2.2 Data Transformation
####2.21 Data Types
We removed the $ in the numbers and changed the data types from character to numeric in each dataset. And we change the column names in the df_api dataset.
```{r}
#salaries by degree
df_deg$start_med_slry <- parse_number(df_deg$start_med_slry)
df_deg$mid_car_slry <- parse_number(df_deg$mid_car_slry)
df_deg$mid_car_10th <- parse_number(df_deg$mid_car_10th)
df_deg$mid_car_25th <- parse_number(df_deg$mid_car_25th)
df_deg$mid_car_75th <- parse_number(df_deg$mid_car_75th)
df_deg$mid_car_90th <- parse_number(df_deg$mid_car_90th)
datatable(df_deg)
```

```{r}
#salaries by college type
df_col$start_med_slry <- parse_number(df_col$start_med_slry)
df_col$mid_car_slry <- parse_number(df_col$mid_car_slry)
df_col$mid_car_10th <- parse_number(df_col$mid_car_10th)
df_col$mid_car_25th <- parse_number(df_col$mid_car_25th)
df_col$mid_car_75th <- parse_number(df_col$mid_car_75th)
df_col$mid_car_90th <- parse_number(df_col$mid_car_90th)
datatable(df_col)
```

```{r}
#salaries by region
df_reg$start_med_slry <- parse_number(df_reg$start_med_slry)
df_reg$mid_car_slry <- parse_number(df_reg$mid_car_slry)
df_reg$mid_car_10th <- parse_number(df_reg$mid_car_10th)
df_reg$mid_car_25th <- parse_number(df_reg$mid_car_25th)
df_reg$mid_car_75th <- parse_number(df_reg$mid_car_75th)
df_reg$mid_car_90th <- parse_number(df_reg$mid_car_90th)
datatable(df_reg)
```

```{r}
names(df_api) <-c("school","median_6yr","median_8yr","median_10yr")
```


#3. Exploratory Data Analysis
##3.1 Median salary analysis in the dataset from EDU
```{r}
df_api_median <- df_api %>%
  select(median_6yr,median_8yr,median_10yr) %>%
  summarise_all(.,funs(median),na.rm = TRUE)
datatable(df_api_median)
```

##3.2 Starting salary vs mid career salaries distribution analysis in the dataset from WSJ
```{r}
start_vs_mid <- df_reg%>%
                 select(start_med_slry, mid_car_slry)%>%
                 gather(career, salary)%>%
                 mutate(career=as_factor(career,fct_rev))
head(start_vs_mid)
```

```{r}
ggplot(start_vs_mid, aes(salary, fill=career))+
  geom_histogram(position ="dodge")
```

Median of starting and mid salary
```{r}
start_med<- start_vs_mid%>%
  filter(career=="start_med_slry")%>%
  summarize(median(salary))
start_med

mid_med<- start_vs_mid%>%
  filter(career=="mid_car_slry")%>%
  summarize(median(salary))
mid_med
```

The distribution for starting median salary is definitely concentrated at the lower range of salaries and is right-skewed. Graduates of most schools start out with median of $45,000. Towards mid-career, the distribution of median salaries becomes more dispersed and the median increases to $80,000.

####3.21 Scatter plot
Analyze the correlation between starting and mid-career salaries.
```{r}
ggplot(df_reg, aes(start_med_slry,mid_car_slry))+
  geom_point()+
  geom_smooth()
```

From the scatter plot above, we can see that there is a strong positive correlaton between the starting salaries and the mid-career salaries.

##3.3 Salary by College Location
```{r}
summary(df_reg$region)
```
```{r}
summary(df_reg$start_med_slry)
```

```{r}
p1<-ggplot(df_reg,aes(x=region,y=start_med_slry))+
         geom_boxplot(aes(fill=region))+
         ggtitle('Starting Median Salaries by Region')+
         xlab('College Location')+
         ylab('Salary ')

p2<-ggplot(df_reg,aes(x=region,y=mid_car_slry))+
         geom_boxplot(aes(fill=region))+
         ggtitle('Mid-Career Median Salaries by Region')+
         xlab('College Location')+
         ylab('Salary ')

multiplot(p1, p2, layout=matrix(c(1, 2), 2, 1, byrow=TRUE))
```

##3.4 Salary by College Type
First,we will take a look at how many types of school and count the numbers of each.
```{r}
ggplot(df_col, aes(school_type)) +
  geom_bar()
```
From the plot above we noticed that most of the schools are state schools. 

####3.41 Starting salary vs mid career salaries analysis by college type
```{r}
df_col_type_slry <- df_col %>%
  select(school_type, start_med_slry, mid_car_slry) %>%
  gather(timeline, salary, start_med_slry:mid_car_slry) %>%
  mutate(timeline = as_factor(timeline, fct_rev))

ggplot(df_col_type_slry, aes(reorder(school_type, salary), salary, fill = timeline)) +
  scale_color_manual(values = c("blue", "pink")) +
  geom_boxplot(alpha = 0.5) +
  scale_fill_manual(values = c("blue", "pink")) +
  theme(legend.position = "right") +
  xlab('School type') 
```

####3.42 Top 10 schools by type
Top 10 schools by median start salary
```{r}
df_start_top10 <- df_col %>%
  select(school_name, school_type, start_med_slry) %>%
  arrange(desc(start_med_slry)) %>%
  top_n(10)

ggplot(df_start_top10, aes(reorder(school_name, start_med_slry), start_med_slry,fill = school_type)) +
  geom_col(alpha = 0.8) +
  geom_text(aes(label = start_med_slry), hjust = 1.1, color = 'gray30') +
  xlab(NULL) +
  coord_flip()
```

Top 10 schools by median mid-career salary
```{r}
df_mid_top10 <- df_col %>%
  select(school_name, school_type, mid_car_slry) %>%
  arrange(desc(mid_car_slry)) %>%
  top_n(10)

ggplot(df_mid_top10, aes(reorder(school_name, mid_car_slry), mid_car_slry,fill = school_type)) +
  geom_col(alpha = 0.8) +
  geom_text(aes(label = mid_car_slry), hjust = 1.1, color = 'gray30') +
  xlab(NULL) +
  coord_flip()
```
In the top 10 list, we can see engineering and ivy league have the highest median salary in both startig and mid-career. 

##3.5 Salary by Majores
####3.51 Starting median salary by major
```{r}
ggplot(df_deg, aes(x = reorder(major, start_med_slry), start_med_slry)) +
  geom_col(fill = "blue", alpha = 0.5) +
  geom_col(aes(x = reorder(major, mid_car_slry), mid_car_slry), alpha = 0.2) +
  geom_text(aes(label = start_med_slry), size = 3, hjust = 1.1) +
  xlab(NULL) +
  coord_flip() +
  ggtitle("Starting salary by major")
```
From the plot above we can see that Physician has the highest median starting salaries. In the top ten highest median starting salaries, most are engineering and computer science related majors. 
   
####3.52 Starting and mid career median salary by major
```{r}
ggplot(df_deg, aes(x = reorder(major, mid_car_slry), mid_car_slry)) +
  geom_col(fill = 'pink',alpha = 0.5) +
  geom_col(aes(x = reorder(major, mid_car_slry), start_med_slry), alpha = 0.4) +
  geom_text(aes(label = mid_car_slry), size = 3, hjust = 1.1) +
  scale_fill_manual(values = c('blue', 'pink')) +
  xlab(NULL) +
  coord_flip() +
  ggtitle("Mid-career salary by major")
```
From the plot above we can see that all of the majors have highest median starting salaries have high salaries in mid-career. 



