---
title: "HW5"
author: "Qixing Li"
date: "March 10, 2019"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE,warning=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### 5.6 Working Backwards, Part II
```{r}
n <- 25

# margin of error is (b-a)/2
ME <- ((77-65)/2)
ME

# sample mean is calculated as (a+b)/2 
se <- ((77+65)/2)
se

# using t(.05)*s/sqrt(n) to calculate the sample standard devation
df <- 25-1
t.value <- qt(.95, df)
t.value
sd <- (ME/t.value)*5
sd
```

### 5.14 SAT Scores

#### a) Raina wants to use a 90% confidence interval. How large of a sample should she collect?
```{r}
me1 <- 25
s.d <- 250
z <- qnorm(0.95)
z
n <- ((z * s.d)/ me1)^2
n
```

#### b) Luke wants to use a 99% confidence interval. Without calculating the actual sample size, determine whether his sample should be larger or smaller than Raina's, and explain your reasoning.
```
Luke's sample size should be larger than Raina's.
```

#### c) Calculate the minimum required sample size for Luke.
```{r}
me2 <- 25
sd <- 250
z <- qnorm(0.99)
z
n <- ((z * sd)/ me2)^2
n
```


### 5.20 High School and Beyond, Part I

#### a) Is there a clear difference in the average reading and writing scores?
```
No. The difference in schores is approximately normal.
```

#### b) Are the reading and writing scores of each student independent of each other?
```
Yes, each student's readinga and writing scores are independent of other student's scores.
```

#### c) Create hypotheses appropriate for the following research question: is there an evident difference in the average score of students in the reading and writing exam?
```
H0: There is no difference in the average of reading and writing score.
Ha: There is difference in the average of reading and writing score.
```

#### d) Check the conditions required to complete this test.
```
Independence of observations and the observations should be aprroximately normal distribution.
```

#### e) The average observed difference in scores iss x{read-write}=-0.545, and the standard deviation of the differences is 8.887 points.
```{r}
mean_diff <- -0.545
sd <- 8.887
n <- 200

# Find the standard erro
sde <- sd / sqrt(n)
# T stat
t <- (mean_diff - 0)/ sde
df1 <- n - 1
p <- pt(t, df=df1)
p
```
```
P is more than 0.05 we fail to reject the null hypothesis. 
```

#### f) What type of error might we have made? Explain what the error means in the context of the application.
```
It is Type II error, we incorrectly reject the alternative hypothesis.
```

#### g) Based on the results of this hypothesis test, would you expect a confidence interval for the average difference between the reading and writing scores to include 0? Explain your reasoning.
```
There is no difference between writing and reading, because we fail to reject null hypothesis, 
```


### 5.32 Fuel efficiency of manual and automatic cares, Part I
H0: There is no difference between automatic car's MPG and manual car's MPG.
Ha: There is differnce between automatic car's MPG and manual car's MPG.

```{r}
n <- 26
mean_diff1 <- 16.12 - 19.85
standard_erro_diff <- sqrt((3.58^2/n)+(4.51^2/n))
standard_erro_diff
t1 <- (mean_diff1 - 0)/ standard_erro_diff
t1 
P_value <- pt(t1, df=n-1)
P_value
```
```
We reject the null hypothesis and accept alternative hypothesis because the p-value is less than 0.05. In this case, there is strong evidence to support the difference between manual and automatic transmissions.
```

### 5.48 Work hours and education

#### a) Write the hypotheses for evaluating whether the average number of hours worked varies across the five groups.
```
H0:There is no difference in average hours worked across the 5 groups.
Ha: At least 1 group has a difference in the average hours worked with the other groups.
```

#### b) Check conditions and describe any assumptions you must make to proceed with the test.
```
The observations are independent and approximately normal, and the variablity across the groups is about the same.
```

#### c) Below is part of the output associated with this test. Fill in the empty cells.
```{r}
n <- 1172
k <- 5
df1 <- k -1
df2 <- n-k
mean_total <- 40.45
group <- data.frame(n=c(121,546,97,253,155),
                    sd=c(15.81,14.97,18.1,13.62,15.51),
                    mean=c(38.67, 39.6, 41.39,42.55,40.85))
sg <- sum (group$n * (group$mean - mean_total)^2)
sg
Msg <- (1/df1)*sg
sse <- 267382
mse <- sse / df2
mse
F <- Msg / mse
F
```
|         |Df   | Sum Sq  |Mean Sq| F value  | Pr (>F) |
|---------|-----|---------|-------|----------|---------|
|degree   | **4**    |  **2004.1** | 501.54  | **2.186745** | 0.0682 |
|Residuals| **1167** |  267382| **229.12**   |              |
|Total    | **1171** | **269386.1** |        |              |


#### d) What is the conclusion of the test?
```
Since P-value = 0.0682, that is greater than 0.05, we fail to reject the null hypothesis. There is no difference between groups.
```





