---
title: "Project Proposal"
author: "Qixing Li"
date: "April 7, 2019"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float:
      collapsed: false
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
wine <- read.csv('wineQualityReds.csv')
```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**
```
Which variables in the dataset are responsible for the quality of the wine. 
```


### Cases 

**What are the cases, and how many are there?**
```
Each case represents one wine. There are 1599 observations in the given data set.
```


### Data collection 

**Describe the method of data collection.**
```
The data set is created by Paulo Cortez (Univ. Minho), Antonio Cerdeira, Fernando Almeida, Telmo Matos and Jose Reis (CVRVV) @ 2009.
```


### Type of study 

**What type of study is this (observational/experiment)?**

```
This is an observational study.
```

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**
```
Data is available online here:
https://www.kaggle.com/piyushgoyal443/red-wine-dataset/downloads/red-wine-dataset.zip/1

P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. Modeling wine preferences by data mining from physicochemical properties. In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.
Available at: http://dx.doi.org/10.1016/j.dss.2009.05.016  
```



### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**
```
The response variable is quality and is numerical.
```

### Explanatory

**What is the explanatory variable, and what type is it (numerical/categorical)?**
```
All the explanatory variablies fixed.acidity, volatile.acidity,citric.acid,residual.sugar,chlorides,free.sulfur.dioxide,total.sulfur.dioxide,density,pH,sulphates,alcohol are numerical. 
```


### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
names(wine)
str(wine)
```

#### Univariate Analysis
**Red wine : Quality**
```{r}
summary(wine$quality)
ggplot(aes(x = quality, color = I('white')), data = wine)+
  geom_bar()+ scale_x_continuous(breaks = seq(3, 8, 1))
```

#### Bivariate Analysis
**Box Plot: Alcohol vs quality**
```{r}
ggplot(aes(x = as.factor(quality), y = wine$alcohol),
  data = wine)+
  geom_boxplot()+
  scale_x_discrete(breaks = seq(1, 10, 1))+
  scale_y_continuous(breaks = seq(8, 15, 0.5))
```
